### 2025-05-05 chapters/3-jets.tex
Pre-correction version of thesis

\begin{figure}[]
    \centering
    % \vstretch{0.3}{% TODO: remove vstretch for renorm illustration
    \includegraphics[width=0.8\textwidth]{example-image-a}
    % \\
    % \includegraphics[width=0.8\textwidth]{example-image-b}
    % \\
    % \includegraphics[width=0.8\textwidth]{example-image-c}
    % }
    \caption[Illustration of renormalization in three contexts (magnetic domains, snowflakes, and jets) which exemplify how zooming into or out of the constituents of a system tends to take the system nearer to a renormalization-group fixed point.]{
        Illustration of renormalization through zooming out of microscopic systems in three contexts:
        %
        (a) iron atoms group to form single and multiple magnetic domains and, eventually, a disordered metallic system;
        \sam{magnet (atom \(\to\) domain \(\to\) domains \(\to\) disorder}
        (b) water molecules form intricate ice structures and snowflake and, eventually, a snow bank;
        %
        \sam{snow (molecule \(\to\) beginning of branch of snowflake \(\to\) snowflake \(\to\) snowbank}
        (c) quarks behave freely at infinite energy, but interact weakly and even form a partonic cascade at collider energies, which themselves are formed by a disordered system of lower-energy hadrons.
        %
        In each case, we encounter a generic phenomenon:
        %
        zooming into or out of the constituents of a system tends to push resulting system closer to a renormalization-group fixed point.
    }
    \label{fig:renormalization}
\end{figure}

@@

One of the first uses of renormalization -- apart from the study of \gls{qcd} itself -- was in the study of the presence and size of magnetic domains within magnetic materials, such as naturally-occurring iron.
%
Renormalization tells us how water molecules form snowflakes, and snowflakes form snow banks.
%
It is the theory of how structure and order forms from simplicity, and how disorder and chaos form from structure.

@@

\begin{definitionbox}{Resummation}{resummation}
    \glslink{resummation}{\vocab{Resummation}} is any method uused to turn (potentially divergent) series expansion into compact and often analytic mathematical expressions without any summation.
    %
    For our purposes, resummation will refer to the process of systematically reintroducing terms that have been previously expanded or neglected in perturbative expansions, in order to account for important contributions at higher orders or at specific scales.
\end{definitionbox}

@@

Finally, we are ready to describe the properties of the entire tree of partonic splittings.
%
As promised, we will compute the probability distribution for the properties of the partonic cascade by sewing together the results of consecutive partonic splittings, leading us to several deep and quite simple conclusions about the fractal nature of jets.
%
The simple description we provide in this chapter strictly holds only in \textit{strongly-ordered} regions of phase space discussed below, and is associated with results at \textit{double logarithmic accuracy} (see \Fig{dlla-plane}).
%
Nonetheless, the model we develop in this chapter facilitates a quantitative study of jet substructure.

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.7\textwidth]{example-image-a}
    \caption[
        Regions of the \(1/z\)-\(s\) plane associated with our approximations.
        %
        Adapted from \Reff{Collins:1984xc} \sam{incomplete}
    ]
    {
        Regions of the \(1/x\)-\(s\) plane associated with the approximations use in this thesis.
        %
        Adapted from \underline{QCD and Collider Physics} \cite{Collins:1984xc}.  \sam{complete -- see page 132, fig 4.18}
    }
    \label{fig:dlla-plane}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\textwidth]{example-image-a}
    \caption[The ripple picture of the partonic cascade \sam{incomplete}]{\sam{The ripple picture of the partonic cascade}}
    \label{fig:}
\end{figure}

@@

Explicitly constructing a simple parton shower is the business of \Prob{parton-shower}.
%
Now, let usmplications of our model.

@@

It may seem miraculous, then, that a detailed analysis of \Eq{series-multiple-emissions} involving Laplace transforms shows that the effects of multiple emissions can \textit{still} be captured only with the radiator for \(X\) -- the integral over the veto region which appears in the \glslink{accuracy}{LL} results above.
%
In particular, the most important effects due to a more detailed analysis of multiple emission contributions to additive observables can be captured by the cumulative distribution
\begin{align}
    \label{eq:resummed-multiple-emissions}
    \Sigma_X(x)
    &=
    \frac{\exp\le(-R_X(x) - \gamma_E R'_X(x)\ri)}{\Gamma\le(1 + R'_X(x)\ri)}
    \,,
\end{align}
where \(\gamma_E=0.577\dots\) is the Euler-Mascheroni constant, \(\Gamma\) is the familiar Euler Gamma function, and \(R'_X(x) := -\dd R_X(x) / \dd \log(x)\).
%
The proof of this formula is the business of \Prob{multiple-emissions}.

@@

\remark{}{
    Conditional distributions of jet parameters can also be important tools in physical applications. In general, the best way to calculate conditional distributions is not to calculate a single Sudakov exponent associated with the related parameters. Instead, it is best to compute the Sudakov exponent associated with the joint cumulative probability of all the involved parameters, find the associated probability distributions, and then divide by the probability distributions you are taking as conditional:
    \begin{align}
        \rho(a | b, c, ...) = \frac{\rho(a, b, c, ...)}{\rho(a, b ...)}
        .
    \end{align}
    %
    Computations of conditional distributions can be important in analytic studies of jet grooming, and make an in the jet grooming calculations of \Chap{grooming}.
}

We have begun to utilize the fractal structure of jets to understand the behavior of jet substructure.
%
By focusing on additive observables, we were able to use the singular features of splittings functions to nonetheless obtain reasonable physical results.
%
Importantly, we were able to leverage the soft singularities and strong angular-ordering of the partonic cascade to achieve two related simplifications, as in \Fig{cascade-simplifications}:

\begin{figure}[t]
    \centering
    \includegraphics[width=0.8\textwidth]{example-image-a}
    \caption[A visualization of the simplifications enabled by additive observables. sam{incomplete}]
    {\sam{simplifications of additive observables in partonic-cascade computations}}
    \label{fig:cascade-simplifications}
\end{figure}

@@

Applying the principles of renormalization to \Eqs{ll_fragmentation_function}{ll_fragmentation_repeat}, we can ask how the process of partonic fragmentation changes as we change the angular resolution of our ``measurements'' (i.e. the jet radius \Rjet{} and the subjet radius \rsub{}).
%
In particular, we can use \Eq{ll_fragmentation_angle} to ask how the fragmentation process changes as a function of \(\Rjet\) or \(\rsub\):
\begin{subequations}
\begin{align}
    \label{eq:fragmentation_dglap_upper}
    \frac{\dd}{\dd\log\le(Q\,\Rjet\ri)} \hat{F}(m;\,\rsub\leftarrow \Rjet)
    &=
    \frac{\alpha_s(Q\,\Rjet)}{\pi}
    \,\,
    \hat{p}(m)
    \,
    \hat{F}(m;\,\rsub\leftarrow \Rjet)
    \\
    \label{eq:fragmentation_dglap_lower}
    \frac{\dd}{\dd\log\le(Q\,\rsub\ri)} \hat{F}(m;\,\rsub\leftarrow \Rjet)
    &=
    -\frac{\alpha_s(Q\,\rsub)}{\pi}
    \,\,
    \hat{p}(m)
    \,
    \hat{F}(m;\,\rsub\leftarrow \Rjet)
    ,
\end{align}
\end{subequations}
where we have suppressed the indices of \(\hat{p}\) and \(\hat{F}\) for brevity.
%
Amazingly, we have recovered the celebrated \glslink{dglapeqn}{Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP) equations} \cite{Gribov:1972ri,Dokshitzer:1977sg,Altarelli:1977zs} for our parton-to-parton fragmentation functions:
\begin{align}
    \frac{\dd}{\dd\log\le(Q \Rjet\ri)}
    F_{f\leftarrow i}(z; \rsub \leftarrow \Rjet)
    =
    \frac{\alpha_s(Q\Rjet)}{\pi}
    \int_0^1
    \dd \frac{\dd y}{y}
    p_{j k}(y)
    F_{j\leftarrow i}\le(\frac{z}{y}; \rsub \leftarrow \Rjet\ri)
    \\
    \frac{\dd}{\dd\log\le(Q \rsub\ri)}
    F(z, \rsub \leftarrow \Rjet)
    =
    -
    \frac{\alpha_s(Q\rsub)}{\pi}
    \int_0^1
    \dd \frac{\dd y}{y}
    p_{j k}(y)
    F_{j\leftarrow i}\le(\frac{z}{y}; \rsub \leftarrow \Rjet\ri)
    \,,
\end{align}
where we have left a sum on \(k\) implicit.
%
More generally, \glspl{dglapeqn} state that the distribution of energy fractions \(z\) of partons inside a high energy composite object \(A\) obey similar integro-differential equations in which \(Q\Rjet\) is replaced by a high energy scale \(E_A\) associated with the energy of the composite object.

@@

    \item
    Repat steps 2. and 3. to continue merging branches of the jet until a single branch remains.
\end{enumerate}
%
This algorithm for angular-ordered, iterative \gls{reclustering} is visualized in \Fig{reclustering}.
%
We will often call the result of angular-ordered iterative \gls{reclustering} a \vocab{jet tree}, meaning a binary tree with angular-ordered branches whose final states are the hadronic final states of a jet.

\begin{figure}[]
    \centering
    \subfloat[]{
        % Subfigure graphic
        \includegraphics[width=.48\textwidth]{example-image-a}
        \label{fig:reclustering}
    }
    \subfloat[]{
        % Subfigure graphic
        \includegraphics[width=.48\textwidth]{example-image-a}
        \label{fig:declustering}
    }
    \caption[Cartoons depicting re-clustering and de-clustering. \sam{incomplete}]{
        \sam{Cartoon depicting reclustering and declustering}
    }
    \label{fig:re-and-de-clustering}
\end{figure}

@@

    Any algorithm which leverages \gls{declustering} is called a \vocab{de-clustering algorithm}.
    %
    The \gls{declustering} procedure is visualized in \Fig{declustering}.
\end{definitionbox}

@@

%\begin{figure}[]
%    \centering
%    \subfloat[]{
%        % Subfigure graphic
%        \includegraphics[width=.33\textwidth]{example-image-a}
%        \label{fig:independent-fragmentation}
%    }
%    \subfloat[]{
%        % Subfigure graphic
%        \includegraphics[width=.33\textwidth]{example-image-b}
%        \label{fig:string-models}
%    }
%    \subfloat[]{
%        % Subfigure graphic
%        \includegraphics[width=.33\textwidth]{example-image-c}
%        \label{fig:cluster-models}
%    }
%    \caption[Cartoons of common hadronization models.]{
%        Cartoons of hadronization as modelled by (a) independent fragmentation, (b) string models, and (c) cluster models.
%        %
%        {todo}
%    }
%    \label{fig:cms-cartoon}
%\end{figure}

@@

        A pioneering hadronization model which played an important role in early studies of jet phenomenology and, for many collider studies, remains a viable tool to this day.
        %
        Also based on the principle of \glslink{lphd}{LPHD}, independent fragmentation models assume that partons \textit{fragment} into hadrons with a \vocab{\gls{parton-to-hadron}} qualitatively similar to the \gls{parton-to-parton} of \Sec{p2p-fragmentation} \cite{Collins:2023cuo,Collins:2011zzd}.
        % https://arxiv.org/pdf/2309.03346
        % and textbook
        %
        The \gls{parton-to-hadron} \(d_{h/j}\) is defined as the number density of final-state hadrons of type \(h\) produced by a parton of type \(j\).
        %
        Like parton distribution functions or splitting functions, \glspl{parton-to-hadron} are portable objects which are nearly universal and may be used in a wide variety of applications;
        %
        they may be defined in terms of operators or extracted phenomenologically.
        %
        However, their theoretical interpretation is somewhat murky -- a single parton has color charge and cannot truly be considered on its own, leading to associated difficulties when using rigorous lattice QFT methods to determine the behavior of fragmentation functions.
        %
        Furthermore, independent fragmentation models are not well suited for capturing detailed multi-particle correlations between final-state hadrons \cite{JADE:1983mvo}.

@@

        String models, the most popular and flexible models of hadronization today, suggest that the energy between quark and anti-quark pairs are contained within narrow tubes, called strings, which then break to form mesons and baryons.
        %
        The presence of \gls{qcd} strings is supported quantitatively by studies in lattice quantum field theory (see e.g. \Reff{Greensite:2011zz}), which are able to probe the properties and dynamics of the \gls{qcd} string directly.
        %
        The Lund string model is the most popular implementation, and is used by \texttt{Pythia} \cite{Bierlich:2022pfr} and therefore all of the phenomenological studies presented in this thesis;
        %
        the Lund string model implemented by \texttt{Pythia} combines the dynamics of \gls{qcd} strings and fragmentation models motivated by quantum tunneling effects (the Schwinger pair production of quarks within the strong gluon field of the \gls{qcd} string) into an extremely flexible and flexible model of hadronization.
        %
        String models of \gls{qcd} were also the birthplace of the field of fundamental string theory.
        %
        It may be no surprise, then, that while string models of hadronization are extremely successful numerical tools, it is unfortunately difficult to obtain analytic results involving hadronization models of \gls{qcd} strings.

@@

One convenient, analytically tractable, independent-fragmentation-like model for non-perturbative effects on jet substructure based on the principle of \glslink{lphd}{LPHD}, is the \vocab{frozen coupling} model, which \textit{freezes} the \gls{qcd} coupling constant below a non-perturbative scale \(\mu_{\rm freeze}\):
\begin{align}
    \alpha^{\rm 1-loop,\,frozen}_s(\kappa; \mu_{\rm freeze})
    =
    \alpha^{\rm 1-loop}_s(\kappa)
    \Theta(\kappa > \mu_{\rm freeze})
    +
    \alpha^{\rm 1-loop}_s(\mu_{\rm freeze})
    \Theta(\mu_{\rm freeze} > \kappa)
    .
    \label{eq:frozen-coupling}
\end{align}
%
\Eq{frozen-coupling} is a simplified analog of several motivated approaches to modeling the infra-red structure of QCD for which \(\alpha_s\) also approximately freezes below a non-perturbative scale.
%
We take the approach of freezing the strong coupling below the scale \(\mu_{\rm freeze}\) = 1 GeV to estimate the effects of hadronization on jet grooming in our numerical calculations of \Sec{pira-resummed}, and we note that changing our choice of \(\mu_{\rm freeze}\) leads only to small effects in substructure distributions.

@@

The \glslink{ue}{underlying event (UE)} is a theoretical source of additive contamination in hadronic collisions whose behavior must be modelled for accurate predictions of jet substructure.
%
The first-principles framework of \gls{ue} is an area of active research, but in this thesis we use the well-understood and highly successful framework of \textit{\glslink{mpi}{multiple parton interactions (MPI)}} in \gls{pythia}.

@@

\remark{mpi}{
    The \gls{mpi} model of \gls{pythia} also combines the multiple-parton-scattering model described above with insights from a theoretical description of all-orders effects in scattering (called Regge-Gribov theory, after its creators \cite{Gribov:1967vfb}) in which ``cut pomerons'', beyond the scope of this thesis, contribute to hadronic cross sections through the production of a large number of low-energy gluons.
    %
    An understanding of the ideas of Regge-Gribov theory in effective field theory is an area of ongoing research which, to my knowledge, is spearheaded by the framework of Glauber \gls{scet}.
}

@@

An important experimental source of additive contamination in particle collisions is \glslink{pileup}{pileup (PU)}.
%
Pileup occurs when radiation from an unrelated collision is detector together with the jets, or other high-energy information, from a collision of interest.

Unfortunately, pileup is unavoidable if we want to learn about the universe.
%
When colliding protons at the LHC, for example, the vast majority of collisions yield ``uninteresting'' results -- that is, results which do not produce the high-energy jets or other high-energy phenomena for which we need collision experiments.
%
In order to see and test interesting physical results, we therefore need to collide a huge number of protons.
%
By applying very strong and precise electromagnetic fields, we are able to line up a huge number of protons into a \textit{bunch};
%
particle physics experiments then accelerate two bunches of protons towards each other so that we can observe the results of many collisions.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problems
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problems}
\makeprob{Regimes and Accuracy {better name}}{accuracy}{

    \begin{enumerate}
        \item
            Show that LO \(\implies\) \(Q \gg \LambdaQCD\).

        \item
            Argue that \(\LambdaQCD\) can be expressed as a non-perturbative scale.

        \item
            Show that LL \(\implies\) scale for the observable...

    \end{enumerate}
}

\makeprob{No Emission Probability}{no-emission}{
    {Problem regarding probability of no emission}

    \begin{enumerate}[label=\roman*)]
        \item
            Between two angles

        \item
            Between two mass scales

         \item
             Between an angle and a mass scale
    \end{enumerate}
}

\makeprob{\(\star\) A Numerical Parton Shower \(\star\)}{parton-shower}{
    \begin{enumerate}[label=\alph*)]
        \item
            {Veto algorithm?}

        \item
            {Code flow}

        \item
            {Write or compare to mine}

        \item
            {Link to manual for mine}
    \end{enumerate}
}
\makeprob{Fragmentation in a Parton Shower}{p2p-fragmentation}{
    {Problem regarding parton shower evolution}

    {Comparison}
    % Based on \url{https://www.sciencedirect.com/science/article/pii/0550321378900159}

    In this problem, we use the following notation:
    \begin{itemize}
        \item
            Let mid-Roman letters (\(i,\,j,\,k\)) denote flavors of partons;

        \item
            Let \(\mc P_{j\leftarrow i}(z) \dd z\) denote the probability of finding a parton of flavor \(j\) with momentum fraction between \(z\) and \(z + \dd z\) of the initiating parton \(i\);

        \item
            Let \(a_s\,\,p_{k_1\,k_2 \leftarrow i}(z) \dd z\) denote the probability that a parton of flavor \(i\) splits exactly into two partons with flavors \(k_1\) (with energy fraction between \(z\) and \(z + \dd z\)) and \(k_2\).
            %
            We will also recall the definition of the inclusive splitting functions, \(\sum_{k'} P_{j\,k' \leftarrow i}(z) =: p_{j\leftarrow i}(z)\).
    \end{itemize}

    We will also make the assumption that \(\alpha_s\), and hence \(a_s\), is independent of the energy scale associated with the splitting.

    \begin{enumerate}[label=\roman*)]
        \item
            Argue on physical grounds that
            \begin{align}
                \label{eq:feynman_field_fragmentation_formula}
                \mc P_{j\leftarrow i}(z)
                &=
                c \delta(1-z) \delta_{ij}
                +
                a_s \, p_{j \leftarrow i}(z)
                +
                a_s \int_z^1 \frac{\dd y'}{y'} \, \sum_{k_1',\,k_2'} \,
                    \mc P_{j\leftarrow k'}(z/y')
                    \,
                    p_{k' \leftarrow i}(y')
                ,
            \end{align}
            and write the associated formula in Mellin space.
            % \begin{align}
            %     \label{eq:feynman_field_fragmentation_formula_mellin}
            %     \hat {\mc P}_{j\leftarrow i}(n)
            %     &=
            %     a_s
            %     \le(
            %         \hat {p}_{j\leftarrow i}(n)
            %         +
            %         \hat {\mc P}_{j\leftarrow k'}(n)
            %         \,\,
            %         \hat {p}_{k' \leftarrow i}(n)
            %     \ri)
            %     ,
            % \end{align}
            % where we have left the sum on \(k'\) implicit.

>>>>

\reusefigure[t!]{picturebook_universe}

@@

One of the first uses of renormalization -- apart from the study of \gls{qft} and \gls{qcd} itself -- was in the study of the presence and size of magnetic domains within magnetic materials, such as naturally-occurring iron.
%
Renormalization tells us how water molecules form snowflakes, and snowflakes form snow banks.
%
It is the theory of how structure and order forms from simplicity, and how disorder and chaos form from structure.

@@

\begin{definitionbox}{Resummation}{resummation}
    \glslink{resummation}{\vocab{Resummation}} is any method used to turn (potentially divergent) series expansion into compact and often analytic mathematical expressions without any summation.
    %
    For our purposes, resummation will refer to the process of systematically reintroducing terms that have been previously expanded or neglected in perturbative expansions, in order to account for important contributions at higher orders or at specific scales.
\end{definitionbox}

@@

Finally, we are ready to describe the properties of the entire tree of partonic splittings.
%
As promised, we will compute the probability distribution for the properties of the partonic cascade by sewing together the results of consecutive partonic splittings, leading us to several deep and quite simple conclusions about the fractal nature of jets.
%
The simple description we provide in this chapter strictly holds only in \textit{strongly-ordered} regions of phase space discussed below, and is associated with results at \textit{double logarithmic accuracy}.
%
Nonetheless, the model we develop in this chapter facilitates a quantitative study of jet substructure.

@@

@@

@@

Explicitly constructing a simple parton shower is the business of \Probs{veto-algorithm}{parton-shower}.
%
Now, let us explore the implications of our model.

@@

It may seem miraculous, then, that a detailed analysis of \Eq{series-multiple-emissions} involving Laplace transforms shows that the effects of multiple emissions can \textit{still} be captured only with the radiator for \(X\) -- the integral over the veto region which appears in the \glslink{accuracy}{LL} results above.
%
In particular, the most important effects due to a more detailed analysis of multiple emission contributions to additive observables can be captured by the cumulative distribution
\begin{align}
    \label{eq:resummed-multiple-emissions}
    \Sigma_X(x)
    &=
    \frac{\exp\le(-R_X(x) - \gamma_E R'_X(x)\ri)}{\Gamma\le(1 + R'_X(x)\ri)}
    \,,
\end{align}
where \(\gamma_E=0.577\dots\) is the Euler-Mascheroni constant, \(\Gamma\) is the familiar Euler Gamma function, and \(R'_X(x) := -\dd R_X(x) / \dd \log(x)\).
%
The proof of this formula, at least for the simple case of angularities, is the business of \Prob{multiple-emissions}.

@@

\remark{}{
    Conditional distributions of jet parameters can also be important tools in physical applications. In general, the best way to calculate conditional distributions is not to calculate a single Sudakov exponent associated with the related parameters. Instead, it is best to compute the Sudakov exponent associated with the joint cumulative probability of all the involved parameters, find the associated probability distributions, and then divide by the probability distributions you are taking as conditional:
    \begin{align}
        \rho(a | b, c, ...) = \frac{\rho(a, b, c, ...)}{\rho(b, c ...)}
        .
    \end{align}
    %
    Computations of conditional distributions can be important in analytic studies of jet grooming, and make an in the jet grooming calculations of \Chap{grooming}.
}

We have begun to utilize the fractal structure of jets to understand the behavior of jet substructure.
%
By focusing on additive observables, we were able to use the singular features of splittings functions to nonetheless obtain reasonable physical results.
%
Importantly, we were able to leverage the soft singularities and strong angular-ordering of the partonic cascade to achieve two related simplifications:

@@

Applying the principles of renormalization to \Eqs{ll_fragmentation_function}{ll_fragmentation_repeat}, we can ask how the process of partonic fragmentation changes as we change the angular resolution of our ``measurements'' (i.e. the jet radius \Rjet{} and the subjet radius \rsub{}).
%
In particular, we can use \Eq{ll_fragmentation_angle} to ask how the fragmentation process changes as a function of \(\Rjet\) or \(\rsub\):
\begin{subequations}
\begin{align}
    \label{eq:fragmentation_dglap_upper}
    \frac{\dd}{\dd\log\le(Q\,\Rjet\ri)} \hat{F}(m;\,\rsub\leftarrow \Rjet)
    &=
    \frac{\alpha_s(Q\,\Rjet)}{\pi}
    \,\,
    \hat{p}(m)
    \,
    \hat{F}(m;\,\rsub\leftarrow \Rjet)
    \\
    \label{eq:fragmentation_dglap_lower}
    \frac{\dd}{\dd\log\le(Q\,\rsub\ri)} \hat{F}(m;\,\rsub\leftarrow \Rjet)
    &=
    -\frac{\alpha_s(Q\,\rsub)}{\pi}
    \,\,
    \hat{p}(m)
    \,
    \hat{F}(m;\,\rsub\leftarrow \Rjet)
    ,
\end{align}
\end{subequations}
where we have suppressed the indices of \(\hat{p}\) and \(\hat{F}\) for brevity.
%
Amazingly, we have recovered the celebrated \glslink{dglapeqn}{Dokshitzer-Gribov-Lipatov-Altarelli-Parisi (DGLAP) equations} \cite{Gribov:1972ri,Dokshitzer:1977sg,Altarelli:1977zs} for our parton-to-parton fragmentation functions:
\begin{subequations}
\label{eq:dglap-equations-parton-shower}
\begin{align}
    \frac{\dd}{\dd\log\le(Q \Rjet\ri)}
    F_{f\leftarrow i}(z; \rsub \leftarrow \Rjet)
    =
    \frac{\alpha_s(Q\Rjet)}{\pi}
    \int_0^1
    \frac{\dd y}{y}
    p_{j k}(y)
    F_{j\leftarrow i}\le(\frac{z}{y}; \rsub \leftarrow \Rjet\ri)
    \\
    \frac{\dd}{\dd\log\le(Q \rsub\ri)}
    F(z, \rsub \leftarrow \Rjet)
    =
    -
    \frac{\alpha_s(Q\rsub)}{\pi}
    \int_0^1
    \frac{\dd y}{y}
    p_{j k}(y)
    F_{j\leftarrow i}\le(\frac{z}{y}; \rsub \leftarrow \Rjet\ri)
    \,,
\end{align}
\end{subequations}
where we have left a sum on \(k\) implicit.
%
More generally, \glspl{dglapeqn} state that the distribution of energy fractions \(z\) of partons inside a high energy composite object \(A\) obey similar integro-differential equations in which \(Q\Rjet\) is replaced by a high energy scale \(E_A\) associated with the energy of the composite object.

@@

    \item
    Repat steps 2. and 3. to continue merging branches of the jet until a single branch remains.
\end{enumerate}
%
We will often call the result of angular-ordered iterative \gls{reclustering} a \vocab{jet tree}, meaning a binary tree with angular-ordered branches whose final states are the hadronic final states of a jet.

@@

    Any algorithm which leverages \gls{declustering} is called a \vocab{de-clustering algorithm}.
\end{definitionbox}

@@

@@

        A pioneering hadronization model which played an important role in early studies of jet phenomenology and, for many collider studies, remains a viable tool to this day.
        %
        Also based on the principle of \glslink{lphd}{LPHD}, independent fragmentation models assume that partons \textit{fragment} into hadrons with a \vocab{\gls{parton-to-hadron}} qualitatively similar to the \gls{parton-to-parton} of \Sec{p2p-fragmentation} \cite{Collins:2023cuo,Collins:2011zzd}.
        %
        The \gls{parton-to-hadron} \(d_{h/j}\) is defined as the number density of final-state hadrons of type \(h\) produced by a parton of type \(j\).
        %
        Like parton distribution functions or splitting functions, \glspl{parton-to-hadron} are portable objects which are nearly universal and may be used in a wide variety of applications;
        %
        they may be defined in terms of operators or extracted phenomenologically.
        %
        However, their theoretical interpretation is somewhat murky -- a single parton has color charge and cannot truly be considered on its own, leading to associated difficulties when using rigorous lattice QFT methods to determine the behavior of fragmentation functions.
        %
        Furthermore, independent fragmentation models are not well suited for capturing detailed multi-particle correlations between final-state hadrons \cite{JADE:1983mvo}.

@@

        String models, the most popular and flexible models of hadronization today, suggest that the energy between quark and anti-quark pairs are contained within narrow tubes, called strings, which then break to form mesons and baryons.
        %
        The presence of \gls{qcd} strings is supported quantitatively by studies in lattice quantum field theory (see e.g. \Reff{Greensite:2011zz}), which are able to probe the properties and dynamics of the \gls{qcd} string directly.
        %
        The Lund string model is the most popular implementation, and is used by \texttt{Pythia} \cite{Bierlich:2022pfr} and therefore all of the phenomenological studies presented in this thesis;
        %
        the Lund string model implemented by \texttt{Pythia} combines the dynamics of \gls{qcd} strings and fragmentation models motivated by quantum tunneling effects (the Schwinger pair production of quarks within the strong gluon field of the \gls{qcd} string) into a flexible model of hadronization with tunable parameters.
        %
        String models of \gls{qcd} were also the birthplace of the field of fundamental string theory.
        %
        It may be no surprise, then, that while string models of hadronization are extremely successful numerical tools, it is unfortunately difficult to obtain analytic results involving hadronization models of \gls{qcd} strings.

@@

One convenient, analytically tractable, independent-fragmentation-like model for non-perturbative effects on jet substructure -- based on the principle of \glslink{lphd}{LPHD} -- is the \vocab{frozen coupling} model, which \textit{freezes} the \gls{qcd} coupling constant below a non-perturbative scale \(\mu_{\rm freeze}\):
\begin{align}
    \alpha^{\rm 1-loop,\,frozen}_s(\kappa; \mu_{\rm freeze})
    =
    \alpha^{\rm 1-loop}_s(\kappa)
    \Theta(\kappa > \mu_{\rm freeze})
    +
    \alpha^{\rm 1-loop}_s(\mu_{\rm freeze})
    \Theta(\mu_{\rm freeze} > \kappa)
    .
    \label{eq:frozen-coupling}
\end{align}
%
\Eq{frozen-coupling} is a simplified analog of several motivated approaches to modeling the infra-red structure of QCD for which \(\alpha_s\) also approximately freezes below a non-perturbative scale.
%
We take the approach of freezing the strong coupling below the scale \(\mu_{\rm freeze}\) = 1 GeV to estimate the effects of hadronization on jet grooming in our numerical calculations of \Sec{pira-resummed}, and we note that changing our choice of \(\mu_{\rm freeze}\) leads only to small effects in substructure distributions.

@@

\begin{figure}[t]
    \centering
    \subfloat[]{
        \includegraphics[width=0.48\textwidth]{figures/misc/ue-jets}
        \label{fig:app-ue}
    }
    \subfloat[]{
        \includegraphics[width=0.48\textwidth]{figures/misc/pu-jet}
        \label{fig:app-pu}
    }
    \caption{Visualizations of important sources of additive contamination in particle collisions:
    %
    \hyperref[fig:app-ue]{(a)}
    \glslink{ue}{the underlying event (UE)}, described theoretically through additional interactions from partons which are not directly involved in a hard scattering process, and
    %
    \hyperref[fig:app-pu]{(b)}
    \glslink{pileup}{pileup (PU)}, or radiation from lower energy scatterings of \textit{additional} protons not directly involved in a primary hard scattering.
    }
    \label{fig:app-additive-contamination}
\end{figure}

The \glslink{ue}{underlying event (UE)} (see \Fig{app-ue}) is a theoretical source of additive contamination in hadronic collisions whose behavior must be modelled for accurate predictions of jet substructure.
%
The first-principles framework of \gls{ue} is an area of active research, but in this thesis we use the well-understood and highly successful framework of \textit{\glslink{mpi}{multiple parton interactions (MPI)}} in \gls{pythia}.

@@

\remark{mpi}{
    The \gls{mpi} model of \gls{pythia} also combines the multiple-parton-scattering model described above with insights from a theoretical description of all-orders effects in scattering (called Regge-Gribov theory, after its creators \cite{Gribov:1967vfb}) in which ``cut pomerons'', beyond the scope of this thesis, contribute to hadronic cross sections through the production of a large number of low-energy gluons.
    %%
    %An understanding of the ideas of Regge-Gribov theory in effective field theory is an area of ongoing research which, to my knowledge, is spearheaded by the framework of Glauber \gls{scet} (see, for example, discussion in Secs. 7.2, 8, and 9, of \Reff{Rothstein:2016bsq}, as well as interesting discussions in \Reffs{Moult:2022lfy,Gao:2024qsg,Gao:2024fyz}).
}

@@

An important experimental source of additive contamination in particle collisions is \glslink{pileup}{pileup (PU)} (see \Fig{app-pu}).
%
Pileup occurs when radiation from an unrelated collision is detector together with the jets, or other high-energy information, from a collision of interest.

Unfortunately, pileup is a necessary feature of modern particle collisions if we want to generate enough collision events to learn about the universe.
%
When colliding protons at the LHC, for example, the vast majority of collisions yield ``uninteresting'' results -- that is, results which do not produce the high-energy jets or other high-energy phenomena for which we need collision experiments.
%
In order to see and test interesting physical results, such as the production of the Higgs boson, we therefore need to collide a huge number of protons.
%
By applying very strong and precise electromagnetic fields, we are able to line up a huge number of protons into a \textit{bunch};
%
particle physics experiments then accelerate two bunches of protons towards each other so that we can observe the results of many collisions.

@@

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Problems
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problems}
% \makeprob{Regimes and Accuracy {better name}}{accuracy}{

%     \begin{enumerate}
%         \item
%             Show that LO \(\implies\) \(Q \gg \LambdaQCD\).

%         \item
%             Argue that \(\LambdaQCD\) can be expressed as a non-perturbative scale.

%         \item
%             Show that LL \(\implies\) scale for the observable...

%     \end{enumerate}
% }

\makeprob{The No-Emission Probability (Sudakov Form Factor)}{no-emission}{
    Let \(\Pi_i(m^2 \leftarrow Q^2)\) indicate the probability that, within a partonic cascade initiated by a parton of flavor \(i\) and virtuality \(Q\), that there is no pairwise emission of partons with a mass larger than \(m\);
    %
    this is sometimes called a \vocab{Sudakov form factor} or a \vocab{no-emission probability}.

    \begin{enumerate}[label=\alph*)]
        \item
        Using the strongly-ordered limit, argue that
        \begin{align}
            \Pi_i(m^2)
            \approx
            \exp[
                -
                \,
                \int \dd z
                \,\,
                4\as
                \,
                \sum_j p_{j \leftarrow i}(z)
                \,
                \int \frac{\dd \theta}{\theta}
                \,\,
                \,
                \Theta(z\theta^2 > m^2)
            ]
            \,,
        \end{align}
        and identify where the strongly-ordered limit appears in your analysis.

        Repeat your analysis to calculate the probability that there is no pairwise emission of partons with an \textit{angularity} (\(z_\text{soft}\,\theta^\varsigma\)) less than \(e_\varsigma\);

        \item
        Explicitly compute the analytic form of the LL no-emission probability for angularities that you derived in integral form in (a).
    \end{enumerate}
}

\makeprob{The Algorithms for Generating Partonic Emissions}{veto-algorithm}{
    When simulating the partonic cascade, the complicated form of the splitting functions and running coupling make analytic results more difficult to obtain even for simple quantities (such as the \gls{parton-to-parton} of \Sec{p2p-fragmentation}).
    %
    To solve more complicated problems, it is common to use numerical parton showers and random number generation to simulate the generation of emissions in the partonic cascade.
    %
    In this problem, we explore the basics of some simple sampling techniques.

    \begin{enumerate}[label=\alph*)]
        \item
        \vocab{Inverse transform sampling} can be used when we have a probability distribution for a random variable \(X\) whose cumulative distribution \(F(X)\) has a known inverse \(F^{-1}(y)\).
        %
        Inverse transform sampling leverages that it is relatively simple to sample a random variable \(U\) which is uniformly distributed from 0 to 1, and relies on the fact that \(F^{-1}(U)\) can be used to obtain the probability distribution of \(X\).

        \textit{Show that \(F^{-1}(U)\) has the same cumulative probability distribution as \(X\) and use this fact to justify inverse transform sampling.}
